On the properties of projectile motion and quadratic air resistance

### 2016-04-11 Paper/main.tex
Review of paper by John

	\section{Introduction}
	Projectile motion is the motion in which an object travels under the constant effect of a force field. The simplest form of projectile motion considers a closed system with no air resistance while a more advanced one might consider multiple gravitational pulls or air resistance. This paper will first describe the ideal trajectory and properties of a projectile parabola and then develop and test an analytical solution to quadratic air resistance.
	\section{Trajectory}
	Consider a Euclidean space $\mathbb{R}^3$ with the three axes $x \comma y$ and $z$. At any time $t$ we want to be able to describe the acceleration, velocity and position of an object with the three vectors $A = (a_x \comma a_y \comma a_z)\comma V = (v_x \comma v_y \comma v_z)$ and $P = (x \comma y \comma z)$. Let $\theta$ be the angle between the $xz$ plane and the initial direction of the object and $\alpha$ the angle between the $x$ axis and a projection of the path on the $xz$ plane. This is visualized in the graph below.
	\begin{center}
		\trimbox{0cm 0cm 0cm -0.6cm} {
			\tdplotsetmaincoords{40}{120} 
			\begin{tikzpicture} [scale = 4, tdplot_main_coords, axis/.style={ -> , black, semithick}, 
			vector/.style={-stealth, black, ultra thick}, 
			vector guide/.style={dashed, black, ultra thick}]

@@

@@

	where $F$ is any force, $m$ is the mass of the object and $a$ is the acceleration of the object we can get an expression for $a$.

@@

	This is the gravitational acceleration which we will denote $g$ to avoid confusion with other accelerations. This vector is the force an object will experience towards the earth. In our considered system the $y$ axis is pointing upwards and the gravitational acceleration is pointing downwards towards the ground. This means the vertical acceleration is $-g$ allowing us to express the total accelerations vector $A$ as follows.

@@

	Here $C$ is a constant of integration and can be written as $v_0 \sin\theta$ due to $\theta$ being the angle between the $xz$ plane and the $y$ axis. This allows us to define the total velocity vector $V$.

@@

	Which is a second degree polynomial proving that projectile motion can be described as a parabola.

@@

	Here $y_{max}$ denotes the highest point the object will reach when thrown at the angle $\theta$ and at the speed $v_0$. This relationship can also be derived using the conservation of energy.

@@

@@

@@

@@

	At any given time $t$ we are moving in the $x$, $y$ and $z$ directions at a velocity $v$ which is a combination of $v_x$, $v_y$ and $v_z$. We can determine the absolute velocity $||v||$ using the Pythagorean theorem for three dimensions.

@@

	For the purpose of this section we will not consider the $z$ axis nor the angle $\alpha$ between the $x$ and $z$ axis. The reasoning behind this is that is would obstruct the results we're trying to conclude. Know that these results are applicable to three dimensions with minimal adjustments including but not limited to multiplying uses of $x$ with $\cos\alpha$.

@@

@@

@@

	and we want to determine $\theta$ such that $P_x = x$ and $P_y = y$ where $P$ is the total position vector and $v_0$ and $t$ is the same for both $P_x$ and $P_y$. Since $P_y$ is the more convoluted expression we solve $P_x$ for $t$ and substitute it into $P_y$.

@@

@@

	Setting the second term as negative yields a negative time when $y_0 > 0$ which is nonsense. We will therefore write it as positive from now on. We can now find the length of a parabola and see how it varies with a changing angle $\theta$. Let $R$ denote the range of the projectile and let it be given by $x(t_{max})$.

@@

	We want to write an algorithm that takes the initial height and velocity and returns the optimal angle. The code tests all possible angles $\theta$ in our equation for $R$ and determines which gives the farthest possible throw. The following pseudo code details the procedure.

@@

		\end{algorithmic}
	\end{algorithm}
	The following two plots are the results of the above algorithm with the initial conditions $v_0 = 10\,m/s$, $h_{min} = 0\,m, h_{max} = 20\,m$ and $v_0 = 25\,m/s$, $h_{min} = 0\,m, h_{max} = 100\,m$ respectively.
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=Starting height (m),
		ylabel=Optimal angle (deg)
		] 
		\addplot[solid] coordinates {(0, 44.9944) (1, 39.9581) (2, 34.9963) (3, 30.0631) (4, 25.1758) (5, 20.4775) (6, 16.3121) (7, 13.0176) (8, 10.6398) (9, 8.9725) (10, 7.7865) (11, 6.9156) (12, 6.2510) (13, 5.7238) (14, 5.2999) (15, 4.9504) (16, 4.6581) (17, 4.4118) (18, 4.1941) (19, 4.0050) (20, 3.8388) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=Starting height (m),
		ylabel=Optimal angle (deg)
		] 
		\addplot[solid] coordinates {(0, 44.9944) (5, 40.6857) (10, 36.3370) (15, 31.8736) (20, 27.2098) (25, 22.2594) (30, 17.0455) (35, 12.0493) (40, 8.3308) (45, 6.1306) (50, 4.8530) (55, 4.0508) (60, 3.5065) (65, 3.1112) (70, 2.8132) (75, 2.5783) (80, 2.3892) (85, 2.2345) (90, 2.1028) (95, 1.9882) (100, 1.8908) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	These graph both follow the same pattern that the distance gained from aiming upwards shrinks as the starting height increases. This is because the travel time cannot be changed all that much when the starting height already contributes so much to the distance.
	\\~\\
	The full code is available at GitHub\cite{code} and mostly contained in \texttt{math.js}. Algorithm \ref{alg:optimalAngle} can be found in the function \texttt{bestAngleFromStartingHeight()}.
	\section{Computational approach to air resistance}\label{seq:numericalAirRes}
	\subsection{Change at every time interval}
	At every time interval $\Delta\,t$ we can apply a force on our object in an angle opposite to $\beta$ proportional to our velocity $v$. While this does not give us a equation to work with we can use Euler's method to get a numerical approximation.

	Let $F_D$ be the force in the opposite direction of that which our object is traveling. At a time interval $\Delta t$ we will add a drag velocity $v_{drag}$ to our total velocity vector $v$.
	\begin{align}\nonumber
		\vec{v}_{new} = \vec{v}_{old} + \vec{v}_{drag}
	\end{align}
	In order to extrapolate $\vec{v}_{drag}$ from $F_D$ we need to consider our objects mass which we will denote $m$. This is due to 
	\begin{align}\nonumber
		\vec{F}=m\vec{a} = m\frac{\Delta\,\vec{v}}{\Delta\,t}\implies \Delta \vec{v} = \frac{\vec{F} \cdot \Delta\,t}{m} = \vec{v}_{drag}
	\end{align}
	This means is that the air resistance is inversely proportional the object's mass. While this might seem counterintuitive at first it is important to remember that a system with a higher mass means that the total energy of the system goes up. It is therefore only logical that our object travels further with higher mass; there is nowhere else for the extra energy to go. One can also consider conservation of momentum in terms of the air mass.
	\\~\\
	$\vec{F}_D$ is given by the quadratic drag equation
	\begin{align}\nonumber
		\vec{F}_D &= -\frac12\rho\,A\,C_D\,|v|^2\hat{v} \implies\\\nonumber
		\vec{v}_{drag} &= \begin{bmatrix}
			-0.5\,\rho\,A\,C_D\,v\,v_x\,m^{-1}\,\Delta\,t\\
			-0.5\,\rho\,A\,C_D\,v\,v_y\,m^{-1}\,\Delta\,t
		\end{bmatrix}
	\end{align}
	where $\rho$ is the air density, $A$ is the frontal area of our projectile, $m$ is the mass of the object and $C_D$ is the coefficient of drag which depends on the shape of our projectile. As $\Delta \rightarrow 0$ our calculations will be more computationally intensive but have greater accuracy.
	\subsection{Algorithm}
	We want a way to numerically apply a force to an object at every time interval and keep track of its velocity and position. The algorithm used can be summarized as the following.
	\begin{algorithm}[H]
		\caption{Algorithm for plotting projectile parabola with air resistance}
		\label{alg:airResistance}
		\begin{algorithmic}
			\Procedure{Air resistance array}{$v_0 ,\theta, \rho, C_D, m, y_0$}
			\State $g \gets $ $9.82$
			\State $x \gets 0$
			\State $y \gets y_0$
			\State $v_x \gets \cos\theta \cdot v_0$
			\State $v_y \gets \sin\theta \cdot v_0$
			\State $\Delta \, t \gets 0.001$;
			\State result $\gets$ [ ]
			\State
			\While {($y > 0$)}
			\State $v \gets \sqrt{v_x^2 + v_y^2}$
			\State $v_x \gets v_x - 0.5\cdot\rho\cdot C_D\cdot v\cdot v_x \cdot m^{-1}\cdot\Delta\,t$
			\State $v_y \gets v_y - 0.5\cdot\rho\cdot C_D\cdot v\cdot v_y \cdot m^{-1}\cdot\Delta\,t$
			\State
			\State $v_y \gets v_y -g \cdot \Delta \, t$
			\State
			\State $x \gets x + v_x \cdot \Delta\,t$
			\State $y \gets y + v_y \cdot \Delta\,t$
			\State
			\State result $\gets$ result $+$ ($x$, $y$)
			\EndWhile
			\State
			\State return result
			\EndProcedure

@@

@@

	Consider the simplest case of air resistance; the one-dimensional falling object. Our object does not travel in a parabola and we have previously described this motion as $y(t) = -\frac12gt^2+y_0$. Knowing the air resistance at any time $t$ we can construct a differential equation describing the motion of our object which is as follows

@@

@@

	$C$ is a constant of integration which is added to $t$ at all times. Let $T = t + C \comma v' = \frac{dv}{dt}$. $G$ can now be written as the following in order to concentrate the unknowns

@@

	For simplicity; let $H = \int \tanh(\sqrt{gk}\,t)\,dt, c = \sqrt{gk}, s = c\,t \implies \frac{ds}{dt} = c \implies dt = \frac{ds}{c}$. The integral is now trivial and we can substitute back $s$ after evaluating it.

@@

		\addplot[solid, dashed] coordinates {(0.000, 4.000) (0.010, 4.000) (0.020, 3.998) (0.030, 3.996) (0.040, 3.992) (0.050, 3.988) (0.060, 3.982) (0.070, 3.976) (0.080, 3.969) (0.090, 3.960) (0.100, 3.951) (0.110, 3.941) (0.120, 3.929) (0.130, 3.917) (0.140, 3.904) (0.150, 3.890) (0.160, 3.874) (0.170, 3.858) (0.180, 3.841) (0.190, 3.823) (0.200, 3.804) (0.210, 3.783) (0.220, 3.762) (0.230, 3.740) (0.240, 3.717) (0.250, 3.693) (0.260, 3.668) (0.270, 3.642) (0.280, 3.615) (0.290, 3.587) (0.300, 3.558) (0.310, 3.528) (0.320, 3.497) (0.330, 3.465) (0.340, 3.432) (0.350, 3.399) (0.360, 3.364) (0.370, 3.328) (0.380, 3.291) (0.390, 3.253) (0.400, 3.214) (0.410, 3.175) (0.420, 3.134) (0.430, 3.092) (0.440, 3.049) (0.450, 3.006) (0.460, 2.961) (0.470, 2.915) (0.480, 2.869) (0.490, 2.821) (0.500, 2.772) (0.510, 2.723) (0.520, 2.672) (0.530, 2.621) (0.540, 2.568) (0.550, 2.515) (0.560, 2.460) (0.570, 2.405) (0.580, 2.348) (0.590, 2.291) (0.600, 2.232) (0.610, 2.173) (0.620, 2.113) (0.630, 2.051) (0.640, 1.989) (0.650, 1.926) (0.660, 1.861) (0.670, 1.796) (0.680, 1.730) (0.690, 1.662) (0.700, 1.594) (0.710, 1.525) (0.720, 1.455) (0.730, 1.383) (0.740, 1.311) (0.750, 1.238) (0.760, 1.164) (0.770, 1.089) (0.780, 1.013) (0.790, 0.936) (0.800, 0.858) (0.810, 0.779) (0.820, 0.699) (0.830, 0.618) (0.840, 0.536) (0.850, 0.453) (0.860, 0.369) (0.870, 0.284) (0.880, 0.198) (0.890, 0.111) (0.900, 0.023) (0.910, -0.066) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	The slow start with a negative second derivative matches our intuition and at roughly $t = 0.8\,s$ the second derivative appears to approach zero. This should be due to the object reaching its terminal velocity which we can calculate using equation \eqref{eq:oneDimAirResTerminalVel}.
	\begin{align}\nonumber
		v_{max} = \sqrt{\frac{g}{k}} = \sqrt{\frac{9.82}{0.432}} = 4.77\,m/s
	\end{align}
	We can approximate the derivative at $t=0.8$ as $\frac{\Delta y}{\Delta t} \approx 4.4 \approx v_{max}$. Equation \eqref{eq:oneDimAirResTMax} also predicts a $t_{max}$ which can be tested.
	\begin{align}\nonumber
		t_{max} = \frac{\cosh^{-1}(\exp(y_0\,k))}{\sqrt{gk}} = \frac{\cosh^{-1}(\exp(1.728))}{\sqrt{4.424224}} \approx 1.17\,s
	\end{align}
	The plot reports $t_{max} \approx 1.17\,s$. Comparing this to a vacuum solution we can use equation \eqref{eq:tMaxFromHeight} which gives
	\begin{align}\nonumber
		t_{max} &= \frac{v_0\sin\theta}{g}+\frac{\sqrt{v_{0}^2\sin^2\theta+2g\,y_0}}{g}\\\nonumber
		&= \frac{\sqrt{2\,g\,y_0}}{g} = \frac{\sqrt{78.56}}{9.82} \approx 0.90\,s
	\end{align}
	This is a major discrepancy which adds credit to the accuracy of equation \eqref{eq:oneDimAirResTMax}.

	We can plot another $y(t)$ where $k$ is much smaller which should result in our object not reaching its terminal velocity in such short time. The following measurements are from a small steel ball with $k = 0.00313$.
	\begin{align}\nonumber
		\begin{cases}\nonumber
			C_D &= 0.47 \\
			m &= 0.022\,kg\\
			y_0 &= 3\,m\\
			v_0 &= 0\,m/s\\
			\rho &= 1.225\,kg/m^3\\
			A &= 0.00023916\,m^2
		\end{cases} \implies k = 0.00313
	\end{align}
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ xlabel=t (s), ylabel=y (m)]
		\addplot[solid] coordinates {(0.000, 3.000) (0.010, 3.000) (0.020, 2.998) (0.030, 2.996) (0.040, 2.992) (0.050, 2.988) (0.060, 2.982) (0.070, 2.976) (0.080, 2.969) (0.090, 2.960) (0.100, 2.951) (0.110, 2.941) (0.120, 2.929) (0.130, 2.917) (0.140, 2.904) (0.150, 2.890) (0.160, 2.874) (0.170, 2.858) (0.180, 2.841) (0.190, 2.823) (0.200, 2.804) (0.210, 2.784) (0.220, 2.762) (0.230, 2.740) (0.240, 2.717) (0.250, 2.693) (0.260, 2.668) (0.270, 2.642) (0.280, 2.615) (0.290, 2.587) (0.300, 2.558) (0.310, 2.528) (0.320, 2.497) (0.330, 2.466) (0.340, 2.433) (0.350, 2.399) (0.360, 2.364) (0.370, 2.328) (0.380, 2.292) (0.390, 2.254) (0.400, 2.215) (0.410, 2.175) (0.420, 2.135) (0.430, 2.093) (0.440, 2.050) (0.450, 2.007) (0.460, 1.962) (0.470, 1.917) (0.480, 1.870) (0.490, 1.823) (0.500, 1.774) (0.510, 1.725) (0.520, 1.674) (0.530, 1.623) (0.540, 1.570) (0.550, 1.517) (0.560, 1.463) (0.570, 1.407) (0.580, 1.351) (0.590, 1.294) (0.600, 1.236) (0.610, 1.176) (0.620, 1.116) (0.630, 1.055) (0.640, 0.993) (0.650, 0.930) (0.660, 0.866) (0.670, 0.801) (0.680, 0.735) (0.690, 0.668) (0.700, 0.600) (0.710, 0.531) (0.720, 0.461) (0.730, 0.391) (0.740, 0.319) (0.750, 0.246) (0.760, 0.172) (0.770, 0.098) (0.780, 0.022) (0.790, -0.055) };

>>>>

	\section{Introduction}
	Projectile motion is the motion in which an object travels under the constant effect of a force field. The simplest form of projectile motion considers a closed system with no air resistance while a more advanced one might consider multiple gravitational pulls or air resistance \todo{Du skulle kunna säga typ friction forces (istället för air resistance) om du vill hålla det mer generellt}. This paper will first describe the ideal trajectory and properties of a projectile parabola\todo{Här introcerar du projectile parabola lite innan du sagt vad det är, men förstår om du vill göra det} and then develop and test an analytical solution to quadratic air resistance \todo{Kanske borde vara typ "and then progressing to models of quadratic air resistance"}.
	\section{Trajectory}
	Consider a Euclidean space $\mathbb{R}^3$ with the three axes $x \comma y$ and $z$. At any time $t$ we want to be able to describe the acceleration, velocity and position of an object with the three vectors $A = (a_x \comma a_y \comma a_z)\comma V = (v_x \comma v_y \comma v_z)$ and $P = (x \comma y \comma z)$. Let $\theta$ be the angle between the $xz$ plane and the initial direction of the object and $\alpha$ the angle between the $x$ axis and a projection of the path on the $xz$ plane. This is visualized in the graph \todo{figure?} below.
	\begin{center}
		\trimbox{0cm 0cm 0cm -0.6cm} {
			\tdplotsetmaincoords{40}{120} 
			\begin{tikzpicture} [scale = 4, tdplot_main_coords, axis/.style={ -> , black, semithick}, 
			vector/.style={-stealth, black, ultra thick}, 
			vector guide/.style={dashed, black, ultra thick}]

@@

	\todo{Den här är ju fett snygg och ger ett bra intryck för papret}

@@

	where $F$ is any force, $m$ is the mass of the object and $a$ is the acceleration of the object \todo{tycker det borde vara komma eller semikolon eller nåt} we can get an expression for $a$.

@@

	This is the gravitational acceleration \todo{kanske borde säga att (eller varför) den är konstant} which we will denote $g$ to avoid confusion with other accelerations. This vector is the force an object will experience towards the earth. In our considered system the $y$ axis is pointing upwards and the gravitational acceleration is pointing downwards towards the ground. This means the vertical acceleration is $-g$ allowing us to express the total accelerations vector $A$ as follows.

@@

	Here $C$ is a constant of integration and can be written as $v_0 \sin\theta$ due to $\theta$ being the angle between the $xz$ plane and the $y$ axis. \todo{Känns lite omotiverat} This allows us to define the total velocity vector $V$.

@@

	Which is a second degree polynomial \todo{Kanske räcker med att sluta meningen här} proving that projectile motion can be described as a parabola.

@@

	Here $y_{max}$ denotes the highest point the object will reach when thrown at the angle $\theta$ and at the speed $v_0$. This relationship can also be derived using the conservation of energy.\todo{Här är mindre rigoröst (du sa liksom F=ma förut), men fattar definitivt varför} \todo{Dock nice att du tar med det också och visade att härledningen kan göras mycket lättare}

@@

	\todo{Du trollar bort lösningen t=0 när du delar med t, men förstår varför}

@@

	\todo{Kanske typ "The the optimal theta for maximizing R may be determined by studying the derivative of R with respect to theta." Så att det blir lite klarare att du vill optimera theta }

@@

	\todo{d R/d theta ka det väl va}

@@

	\todo{Är det här inte angle at a given time?}
	At any given time $t$ we are moving in the $x$, $y$ and $z$ directions at a velocity $v$ which is a combination of $v_x$, $v_y$ and $v_z$. We can determine the absolute velocity $||v||$ using the Pythagorean theorem for three dimensions.\todo{Här nämner du pyth men inte när du räknar ut R}

@@

	For the purpose of this section we will not consider the $z$ axis nor the angle $\alpha$ between the $x$ and $z$ axis. The reasoning behind this is that is would obstruct \todo{Skulle det verkligen obstruct? Blir det inte bara svårare att se? typ introduce unnecessary complexity } the results we're trying to conclude. Know that these results are applicable \todo{kanske expandable eller may be expanded} to three dimensions with minimal adjustments \todo{Kanske for example istället} including but not limited to multiplying uses of $x$ with $\cos\alpha$.

@@

	\todo{På vilket sätt specifyar du projectile range i den här sektionen?}

@@

	\todo{Känns lite konstigt att theta och v0 är funktioner av konstanter. Känns som att man kan använda funktionerna ovan för att räkna ut ex theta vid en viss tidpunkt, men du vill ju bara att man ska kunna räkna ut theta om man har R och v0 (antar jag)}

@@

	\todo{Ska det inte bara vara x och y? Du har definierat P = (x, y). Du skriver ju ändå om till x och y sen}
	and we want to determine $\theta$ such that $P_x = x$ and $P_y = y$ where $P$ is the total position vector and $v_0$ and $t$ is the same for both $P_x$ and $P_y$ \todo{Behöver du berätta det här?}. Since $P_y$ is the more convoluted expression we solve $P_x$ for $t$ and substitute it into $P_y$.

@@

	\todo{Behöver du definiera pq?}

@@

	Setting the second term as negative yields a negative time when $y_0 > 0$ which is nonsense \todo{Which is not defined eller physically impossible? (nonsense är bra annars)}. We will therefore write it as positive from now on. \todo{kan bindas ihop med förra "... and can/may now find the length"}We can now find the length of a parabola \todo{observing how it varies with...?} and see how it varies with a changing angle $\theta$. Let $R$ denote the range of the projectile and let it be given by $x(t_{max})$. \todo{Kanske bara " given by x(tmax)"}

@@

	\todo{Kanske borde säga varför det ska göras numeriskt: because this can not be solved by analytical methods, it must be solved numerically. For this solution we want to write an algorithm...}
	We want to write an algorithm that takes the initial height and velocity and returns the optimal angle. The code tests all possible angles $\theta$ \todo{Du kan ju inte testa alla} in our equation for $R$ and determines which gives the farthest possible throw. The following pseudo code details the procedure.\todo{The procedure is described in the following pseudo code (kanske)}

@@

		\end{algorithmic}
	\end{algorithm}
	The following two plots are the results of the above algorithm with the initial conditions $v_0 = 10\,m/s$, $h_{min} = 0\,m, h_{max} = 20\,m$ and $v_0 = 25\,m/s$, $h_{min} = 0\,m, h_{max} = 100\,m$ respectively.\todo{Kanske inte behöver säga skalan eftersom det framgår i figuren. Kanske bara typ "is shown in the following two plots"}
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=Starting height (m),
		ylabel=Optimal angle (deg)
		] 
		\addplot[solid] coordinates {(0, 44.9944) (1, 39.9581) (2, 34.9963) (3, 30.0631) (4, 25.1758) (5, 20.4775) (6, 16.3121) (7, 13.0176) (8, 10.6398) (9, 8.9725) (10, 7.7865) (11, 6.9156) (12, 6.2510) (13, 5.7238) (14, 5.2999) (15, 4.9504) (16, 4.6581) (17, 4.4118) (18, 4.1941) (19, 4.0050) (20, 3.8388) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	\begin{center}
		\begin{tikzpicture}
		\begin{axis}[ 
		xlabel=Starting height (m),
		ylabel=Optimal angle (deg)
		] 
		\addplot[solid] coordinates {(0, 44.9944) (5, 40.6857) (10, 36.3370) (15, 31.8736) (20, 27.2098) (25, 22.2594) (30, 17.0455) (35, 12.0493) (40, 8.3308) (45, 6.1306) (50, 4.8530) (55, 4.0508) (60, 3.5065) (65, 3.1112) (70, 2.8132) (75, 2.5783) (80, 2.3892) (85, 2.2345) (90, 2.1028) (95, 1.9882) (100, 1.8908) };
		\end{axis}
		\end{tikzpicture}
	\end{center}
	These graph both follow the same pattern that the distance gained from aiming upwards shrinks as the starting height increases. \todo{Du menar derivatan eller? Kanske borde säga att det är så man ser det} This is because the travel time cannot be changed all that much when the starting height already contributes so much to the distance.
	\\~\\
	The full code is available at GitHub\cite{code} and mostly contained in \texttt{math.js}. Algorithm \ref{alg:optimalAngle} can be found in the function \texttt{bestAngleFromStartingHeight()}.
	\section{Computational approach to air resistance}\label{seq:numericalAirRes}
	\subsection{Change at every time interval}
	At every time interval $\Delta\,t$ we can apply a force on our object in an angle opposite to $\beta$ proportional to our velocity $v$. While this does not give us a equation to work with we can use Euler's method to get a numerical approximation. \todo{Kanske berätta hur luftmotståndsmodellen funkar först (att det är motsat riktningen och = ky), känns lite omotiverat i den här meningen varför man skulle vilja göra det}

	Let $F_D$ be the force in the opposite direction of that which our object is traveling. At a time interval $\Delta t$ we will add a drag velocity $v_{drag}$ to our total velocity vector $v$.
	\begin{align}\nonumber
		\vec{v}_{new} = \vec{v}_{old} + \vec{v}_{drag}
	\end{align}
	In order to extrapolate \todo{Är det verkligen extrapolate} $\vec{v}_{drag}$ from $F_D$ we need to consider our objects mass which we will denote $m$. This is due to 
	\begin{align}\nonumber
		\vec{F}=m\vec{a} = m\frac{\Delta\,\vec{v}}{\Delta\,t}\implies \Delta \vec{v} = \frac{\vec{F} \cdot \Delta\,t}{m} = \vec{v}_{drag}
	\end{align}
	This means is that the air resistance is inversely proportional the object's mass. While this might seem counterintuitive at first it is important to remember that a system with a higher mass means that the total energy of the system goes up \todo{of the system increases}. It is therefore only \todo{Tycker inte only behövs} logical that our object travels further with higher mass; there is nowhere else for the extra energy to go. One can also consider conservation of momentum in terms of the air mass.
	\\~\\
	$\vec{F}_D$ is given by the quadratic drag equation
	\begin{align}\nonumber
		\vec{F}_D &= -\frac12\rho\,A\,C_D\,|v|^2\hat{v} \implies\\\nonumber
		\vec{v}_{drag} &= \begin{bmatrix}
			-0.5\,\rho\,A\,C_D\,v\,v_x\,m^{-1}\,\Delta\,t\\
			-0.5\,\rho\,A\,C_D\,v\,v_y\,m^{-1}\,\Delta\,t
		\end{bmatrix}
	\end{align}
	\todo{Känns som det händer rätt mycket i det där steget}
	where $\rho$ is the air density, $A$ is the frontal area of our projectile, $m$ is the mass of the object and $C_D$ is the coefficient of drag which depends on the shape of our projectile. As $\Delta \rightarrow 0$ our calculations will be more computationally intensive but have greater accuracy.
	\subsection{Algorithm}
	We want a way to numerically apply a force to an object at every time interval and keep track of its velocity and position. The algorithm used can be summarized as the following.
	\begin{algorithm}[H]
		\caption{Algorithm for plotting projectile parabola with air resistance}
		\label{alg:airResistance}
		\begin{algorithmic}
			\Procedure{Air resistance array}{$v_0 ,\theta, \rho, C_D, m, y_0$}
			\State $g \gets $ $9.82$
			\State $x \gets 0$
			\State $y \gets y_0$
			\State $v_x \gets \cos\theta \cdot v_0$
			\State $v_y \gets \sin\theta \cdot v_0$
			\State $\Delta \, t \gets 0.001$;
			\State result $\gets$ [ ]
			\State
			\While {($y > 0$)}
			\State $v \gets \sqrt{v_x^2 + v_y^2}$
			\State $v_x \gets v_x - 0.5\cdot\rho\cdot C_D\cdot v\cdot v_x \cdot m^{-1}\cdot\Delta\,t$
			\State $v_y \gets v_y - 0.5\cdot\rho\cdot C_D\cdot v\cdot v_y \cdot m^{-1}\cdot\Delta\,t$
			\State
			\State $v_y \gets v_y -g \cdot \Delta \, t$
			\State
			\State $x \gets x + v_x \cdot \Delta\,t$
			\State $y \gets y + v_y \cdot \Delta\,t$
			\State
			\State result $\gets$ result $+$ ($x$, $y$)
			\EndWhile
			\State
			\State return result
			\EndProcedure

@@

	\todo{Kanske är lite jobbigt, men du borde kanske ha typ en fil per plot som man kan kolla på som läsare}

@@

	\todo{Du börjar bli mindre rigorös, men att det är jobbigt 8och möjligtvis onödigt att vara det hela tiden}
	Consider the simplest case of air resistance; the one-dimensional falling object. Our object does not travel in a parabola \todo{Du skulle kunna säga att det är för att x(t) = 0} and we have previously described this motion as $y(t) = -\frac12gt^2+y_0$. Knowing the air resistance at any time $t$ we can construct a differential equation describing the motion of our object which is as follows \todo{Tycker inte which is utan bara as follows}

@@

	\todo{Ska det inte vara g/m?}

@@

	$C$ is a constant of integration which is added to $t$ at all times. Let $T = t + C \comma v' = \frac{dv}{dt}$. $G$ can now be written as the following in order to concentrate the unknowns \todo{Aldrig hört concentrate the unknowns formuleringen antar att det betyder att man vill samla dem?}

@@

	For simplicity; let $H = \int \tanh(\sqrt{gk}\,t)\,dt, c = \sqrt{gk}, s = c\,t \implies \frac{ds}{dt} = c \implies dt = \frac{ds}{c}$. The integral is now trivial and we can substitute back $s$ after evaluating it.\todo{Istället för det här resonemanget skulle du bara kunna hänvisa till kedjeregeln och därför dela med sqrt(gk). integral(tanh(x)dx) = ln(cosh(x)) => integral(tanh(kx)dx) = ln(cosh(kx))/k}

@@

		\addplot[solid, dashed] coordinates {(0.000, 4.000) (0.010, 4.000) (0.020, 3.998) (0.030, 3.996) (0.040, 3.992) (0.050, 3.988) (0.060, 3.982) (0.070, 3.976) (0.080, 3.969) (0.090, 3.960) (0.100, 3.951) (0.110, 3.941) (0.120, 3.929) (0.130, 3.917) (0.140, 3.904) (0.150, 3.890) (0.160, 3.874) (0.170, 3.858) (0.180, 3.841) (0.190, 3.823) (0.200, 3.804) (0.210, 3.783) (0.220, 3.762) (0.230, 3.740) (0.240, 3.717) (0.250, 3.693) (0.260, 3.668) (0.270, 3.642) (0.280, 3.615) (0.290, 3.587) (0.300, 3.558) (0.310, 3.528) (0.320, 3.497) (0.330, 3.465) (0.340, 3.432) (0.350, 3.399) (0.360, 3.364) (0.370, 3.328) (0.380, 3.291) (0.390, 3.253) (0.400, 3.214) (0.410, 3.175) (0.420, 3.134) (0.430, 3.092) (0.440, 3.049) (0.450, 3.006) (0.460, 2.961) (0.470, 2.915) (0.480, 2.869) (0.490, 2.821) (0.500, 2.772) (0.510, 2.723) (0.520, 2.672) (0.530, 2.621) (0.540, 2.568) (0.550, 2.515) (0.560, 2.460) (0.570, 2.405) (0.580, 2.348) (0.590, 2.291) (0.600, 2.232) (0.610, 2.173) (0.620, 2.113) (0.630, 2.051) (0.640, 1.989) (0.650, 1.926) (0.660, 1.861) (0.670, 1.796) (0.680, 1.730) (0.690, 1.662) (0.700, 1.594) (0.710, 1.525) (0.720, 1.455) (0.730, 1.383) (0.740, 1.311) (0.750, 1.238) (0.760, 1.164) (0.770, 1.089) (0.780, 1.013) (0.790, 0.936) (0.800, 0.858) (0.810, 0.779) (0.820, 0.699) (0.830, 0.618) (0.840, 0.536) (0.850, 0.453) (0.860, 0.369) (0.870, 0.284) (0.880, 0.198) (0.890, 0.111) (0.900, 0.023) (0.910, -0.066) };
	The slow start with a negative second derivative matches our intuition and at roughly $t = 0.8\,s$ the second derivative appears to approach zero. This should be due to the object reaching its terminal velocity which we can calculate using equation \eqref{eq:oneDimAirResTerminalVel}.
		v_{max} = \sqrt{\frac{g}{k}} = \sqrt{\frac{9.82}{0.432}} = 4.77\,m/s
	We can approximate the derivative at $t=0.8$ as $\frac{\Delta y}{\Delta t} \approx 4.4 \approx v_{max}$. Equation \eqref{eq:oneDimAirResTMax} also predicts a $t_{max}$ which can be tested.
		t_{max} = \frac{\cosh^{-1}(\exp(y_0\,k))}{\sqrt{gk}} = \frac{\cosh^{-1}(\exp(1.728))}{\sqrt{4.424224}} \approx 1.17\,s
	The plot reports $t_{max} \approx 1.17\,s$. Comparing this to a vacuum solution we can use equation \eqref{eq:tMaxFromHeight} which gives
		t_{max} &= \frac{v_0\sin\theta}{g}+\frac{\sqrt{v_{0}^2\sin^2\theta+2g\,y_0}}{g}\\\nonumber
		&= \frac{\sqrt{2\,g\,y_0}}{g} = \frac{\sqrt{78.56}}{9.82} \approx 0.90\,s
	This is a major discrepancy which adds credit to the accuracy of equation \eqref{eq:oneDimAirResTMax}.
	\todo{Fast det säger ju ingenting förren du jämfört med verjligheten}
	We can plot another $y(t)$ where $k$ is much smaller which should result in our object not reaching its terminal velocity in such short time. The following measurements are from a small steel ball with $k = 0.00313$.
			m &= 0.022\,kg\\
			y_0 &= 3\,m\\
			A &= 0.00023916\,m^2
		\end{cases} \implies k = 0.00313
		\addplot[solid] coordinates {(0.000, 3.000) (0.010, 3.000) (0.020, 2.998) (0.030, 2.996) (0.040, 2.992) (0.050, 2.988) (0.060, 2.982) (0.070, 2.976) (0.080, 2.969) (0.090, 2.960) (0.100, 2.951) (0.110, 2.941) (0.120, 2.929) (0.130, 2.917) (0.140, 2.904) (0.150, 2.890) (0.160, 2.874) (0.170, 2.858) (0.180, 2.841) (0.190, 2.823) (0.200, 2.804) (0.210, 2.784) (0.220, 2.762) (0.230, 2.740) (0.240, 2.717) (0.250, 2.693) (0.260, 2.668) (0.270, 2.642) (0.280, 2.615) (0.290, 2.587) (0.300, 2.558) (0.310, 2.528) (0.320, 2.497) (0.330, 2.466) (0.340, 2.433) (0.350, 2.399) (0.360, 2.364) (0.370, 2.328) (0.380, 2.292) (0.390, 2.254) (0.400, 2.215) (0.410, 2.175) (0.420, 2.135) (0.430, 2.093) (0.440, 2.050) (0.450, 2.007) (0.460, 1.962) (0.470, 1.917) (0.480, 1.870) (0.490, 1.823) (0.500, 1.774) (0.510, 1.725) (0.520, 1.674) (0.530, 1.623) (0.540, 1.570) (0.550, 1.517) (0.560, 1.463) (0.570, 1.407) (0.580, 1.351) (0.590, 1.294) (0.600, 1.236) (0.610, 1.176) (0.620, 1.116) (0.630, 1.055) (0.640, 0.993) (0.650, 0.930) (0.660, 0.866) (0.670, 0.801) (0.680, 0.735) (0.690, 0.668) (0.700, 0.600) (0.710, 0.531) (0.720, 0.461) (0.730, 0.391) (0.740, 0.319) (0.750, 0.246) (0.760, 0.172) (0.770, 0.098) (0.780, 0.022) (0.790, -0.055) };
